Extended Abstract

### 2021-07-27 paper/paper.tex
Finish with minor review changes

The \picalc{} \cite{MilnerPW92a,Milner99} models concurrent processing, boiling it down to the transmission of data over communication channels --- where channels too are sent as payload.
The linear \picalc{} \cite{KobayashiPT96} introduces a type system that ensures that every channel is used \emph{exactly} once.
This restriction ensures communication privacy, communication safety, and the absence of race conditions.
We generalise to a \picalc{} with linear and shared types, where the input and output multiplicities of a channel are either $0$ (cannot be used to transmit), $1$ (must be used exactly once), or $\omega$ (unlimited use).

@@

We define a standard syntax for the \picalc{} using type-level de Bruijn indices.
Variable references $i_n$ are of type $\Fin{n}$, expressions $e_n$ and $f_n$ are of type $\sExpr{n}$, processes $p_n$ and $q_n$ are of type $\sProc{n}$.
\[
\begin{aligned}[c]
  e_n ~ f_n  :=
  &~ \sunit \\
  |&~ \svar~i_n \\
  |&~ \spair~e_n~f_n \\
  |&~ \sfst~e_n ~|~  \ssnd~e_n \\
  |&~ \sinl~e_n ~|~  \sinr~e_n
\end{aligned}
\begin{aligned}[c]
  p_n ~ q_n  :=
  &~ \send ~|~  \srec~p_n ~|~ \snew~p_{1+n} \\
  |&~ \srecv~e_n~p_{1+n} \\
  |&~ \ssend~e_n~f_n~p_n \\
  |&~ \scomp~p_n~q_n \\
  |&~ \scase~e_n~p_{1+n}~q_{1+n} \\
\end{aligned}
\]

@@

\paragraph{Typing Judgments}
\label{typing-judgments}
We can finally define our typing judgment for variables, expressions and processes (some cases are omitted for brevity).

@@

Co-contextual typing inference traverses processes bottom-up keeping a typing context and collecting constraints.
This makes inference subproblems independent and thus easily parallelisable.
In the shared \picalc{}, inference constraints always have a most general unifier, in the shared \emph{and} linear \picalc{} however they do not.
Consider the open process $\ssend~a~\sunit~(\ssend~x~a~\send)$ where $x$ and $a$ are free: we partly use $a$ to send, then send $a$ away over $x$ and terminate.
When the process terminates, all multiplicities must be unrestricted, but we cannot decide whether to instantiate with $\tzero$ or $\tomega$.
Similarly, we ignore \emph{how much} of $a$ we are sending away over $x$: that depends on what the receiving process expects.
Nonetheless, we do have to keep track of the fact that $a$ has been used once to send.

>>>>

The \picalc{} \cite{MilnerPW92a,Milner99} models concurrent processing, boiling it down to the transmission of data over communication channels --- where channels too are sent as payload.
The linear \picalc{} \cite{KobayashiPT96} introduces a type system that ensures that every channel is used \emph{exactly} once.
This restriction ensures communication safety, and the absence of race conditions.
We generalise to a \picalc{} with linear and shared types, where the input and output capabilities of a channel are either $0$ (cannot be used to transmit), $1$ (must be used exactly once), or $\omega$ (unlimited use).

@@

We define a standard syntax for the \picalc{} using type-level de Bruijn indices.
Variable references $i_{1+n}$ are of type $\Fin{(1 + n)}$ with constructors $\fzero$ and $\fsuc~i_n$, expressions $e_n$ and $f_n$ are of type $\sExpr{n}$, processes $p_n$ and $q_n$ are of type $\sProc{n}$.
\[
\begin{aligned}[c]
  e_n ~ f_n  :=
  &~ \sunit \\
  |&~ \svar~i_n \\
  |&~ \spair~e_n~f_n \\
  |&~ \sfst~e_n ~|~  \ssnd~e_n \\
  |&~ \sinl~e_n ~|~  \sinr~e_n
\end{aligned}
\begin{aligned}[c]
  p_n ~ q_n  :=
  &~ \send ~|~  \srec~p_n ~|~ \snew~p_{1+n} \\
  |&~ \srecv~e_n~p_{1+n} \\
  |&~ \ssend~e_n~f_n~p_n \\
  |&~ \scomp~p_n~q_n \\
  |&~ \scase~e_n~p_{1+n}~q_{1+n} \\
\end{aligned}
\]

@@

\paragraph{Typing Judgments}
\label{typing-judgments}
We can now define our typing judgment for variables ($\tVar{\Gamma}{i}{t}$, where variable $i$ under context $\Gamma$ is of type $t$), expressions ($\tExpr{\Gamma}{e}{t}$, where expression $e$ under context $\Gamma$ is well typed with type $t$) and processes ($\tProc{\Gamma}{p}$, where process $p$ is well typed under context $\Gamma$). Some cases are omitted for brevity.

@@

Co-contextual typing inference traverses processes bottom-up keeping a typing context and collecting constraints.
This makes inference subproblems independent and thus easily parallelisable.
In the shared \picalc{}, inference constraints always have a most general unifier, in the shared \emph{and} linear \picalc{} however they do not.
Consider the open process $\ssend~a~\sunit~(\ssend~x~a~\send)$ where $x$ and $a$ are free: we partly use $a$ to send, then send $a$ away over $x$ and terminate.
When the process terminates, all usage annotations must be unrestricted, but we cannot decide whether to instantiate with $\tzero$ or $\tomega$.
Similarly, we ignore \emph{how much} of $a$ we are sending away over $x$: that depends on what the receiving process expects.
Nonetheless, we do have to keep track of the fact that $a$ has been used once to send.
